---
title: '*juede* and *zhidao* corpus study'
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    toc: true
--- 

```{r libraries, include=FALSE} 
# load the library
library(childesr)
library(dplyr)
library(lme4)
library(emmeans)
library(tidyverse)
library(ggplot2)
library(ggrepel)
library(ggsignif)
library(ggpattern)
library(tidytext)
library(RColorBrewer)
library(viridis)
library(stringr)
library(irr)
```


```{r setup, include=FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
sentencePalette <- c("#BEBADA", "#80B1D3") 
embedPalette <- c("#BEBADA", "#8DD3C7") 
embedfullPalette <- c("#E41A1C","#4DAF4A","#FFFF33")
```

# 1. Load Data
```{r}
data <- read.csv("../../data/childes/zhidao_juede_no_readings_combined.csv", header=TRUE) %>% 
  filter(!grepl("Chang2/BookReading", file_name)) %>%
  select(!c("file_name", "speaker", "sentence"))

clean_data <- data %>% 
  mutate(id=row_number(),
         across(c(sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3, embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3),
                ~ if_else(.x == "3 how/why", "2 wh_question", .x)),
         across(c(embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3),
                ~ if_else(.x == "3 base-generated wh/polar", "3 base_position", .x)),
         across(c(subject_j1, subject_j2),
                ~ case_when(.x == "4 self_reference" ~ "1 first_person",
                            .x == "5 自己" ~ "5 ambiguous ziji",
                            TRUE ~ .x)))
```

# 2. Reliability 
## 2.1 Type of matrix subject
```{r warning=FALSE}
matrix_subject_reliability.data <- clean_data %>% 
  select(c(subject_j1, subject_j2, id)) %>% 
  # mutate(subject_j2 = if_else(subject_j2 == "", NA, subject_j2)) %>% # a more lenient criterion when change to NA 
  pivot_longer(cols = c(subject_j1, subject_j2),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id) 

matrix_subject_reliability.data <- as.matrix(matrix_subject_reliability.data)
kripp.alpha(matrix_subject_reliability.data, method = "nominal")
```

## 2.2 Type of matrix sentence 
```{r warning=FALSE}
matrix_sentence_reliability.data <- clean_data %>% 
  select(c(sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3, id)) %>% 
  pivot_longer(cols = c(sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id) 

matrix_sentence_reliability.data <- as.matrix(matrix_sentence_reliability.data)
kripp.alpha(matrix_sentence_reliability.data, method = "nominal")
```

## 2.3 Type of embedded sentence 
```{r warning=FALSE}
embedded_sentence_reliability.data <- clean_data %>% 
  select(c(embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3, id)) %>% 
  mutate(
    across(embedded_sentence_type_e:embedded_sentence_type_j3, ~ if_else(.x == "", "NONE", .x))
  ) %>% 
  pivot_longer(cols = c(embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id) 

embedded_sentence_reliability.data <- as.matrix(embedded_sentence_reliability.data)
kripp.alpha(embedded_sentence_reliability.data, method = "nominal")
```

## 2.4 Whether there is an embedded CP
Two annotators coded whether there is an CP or not.
```{r warning=FALSE}
embedded_CP_reliability.data <- clean_data %>% 
  select(c(embedded_CP_j1, embedded_CP_j2, id)) %>% 
  pivot_longer(cols = c(embedded_CP_j1, embedded_CP_j2),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id)

embedded_CP_reliability.data <- as.matrix(embedded_CP_reliability.data)
kripp.alpha(embedded_CP_reliability.data, method = "nominal")
```

There is inconsistency in terms of how annotators treat cases where the verb is followed by an adjective. Here, we explore a more generous way to calculate the reliability, where we combine cases where the annotators indicated that 1) "2 (Yes) CP" for whether there is an embedded CP and 2) "0 No CP" for whether there is an embedded CP AND "1 Adj" for the type of embedded clause to "2 CP". 
```{r warning=FALSE}

embedded_CP_reliability_alt.data <- clean_data %>%
  mutate(embedded_CP_j1 = if_else(embedded_CP_j1 == "0 No" & embedded_content_j1 == "1 Adj", "2 CP", embedded_CP_j1),
         embedded_CP_j2 = if_else(embedded_CP_j2 == "0 No" & embedded_content_j2 == "1 Adj", "2 CP", embedded_CP_j2)) %>% 
  select(c(embedded_CP_j1, embedded_CP_j2, id)) %>% 
  pivot_longer(cols = c(embedded_CP_j1, embedded_CP_j2),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id)

embedded_CP_reliability_alt.data <- as.matrix(embedded_CP_reliability_alt.data)
kripp.alpha(embedded_CP_reliability_alt.data, method = "nominal")
```

## 2.5 negation 
```{r warning=FALSE}
negation_reliability.data <- clean_data %>% 
  select(c(matrix_negation_e, matrix_negation_g, matrix_negation_j1, id)) %>% 
  pivot_longer(cols = c(matrix_negation_e, matrix_negation_g, matrix_negation_j1),
               names_to = "coder_id",
               values_to = "value") %>% 
  pivot_wider(names_from = id,
              values_from = value) %>% 
  select(-coder_id)

negation_reliability.data <- as.matrix(negation_reliability.data)
kripp.alpha(negation_reliability.data, method = "nominal")
```

# 3. Distrubtions
For the distribution, we are only using data that the two coders agreed on.

## 3.1 matrix subject
```{r matrix subject summary, message=FALSE}
matrix_subject.data <- clean_data %>% 
  select(c(id, verb, subject_j1, subject_j2)) %>% 
  mutate(agreement = if_else(subject_j1 == subject_j2, "same", "diff")) %>%
  filter(agreement == "same") %>% 
  select(!c(agreement, subject_j1)) %>% 
  rename(subject_type = subject_j2)

matrix_subject_summary <- matrix_subject.data %>% 
  group_by(verb, subject_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

```{r matrix subject plot}
ggplot(matrix_subject_summary, 
       aes(x=verb,
           y=percent,
           fill=subject_type)) + 
  geom_col()+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  # geom_label(aes(label=count,
  #                group = subject_type),
  #           position = position_stack(vjust = 0.5),
  #           fill = "white",
  #           color = "black",
  #           label.size = 0.3,
  #           label.padding = unit(0.15, "lines"),
  #           size = 3)+
  theme_bw()+
  scale_fill_brewer(palette = "Set2",
                    name="Matrix subject type",
                    guide=guide_legend(nrow=2)) +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```

## 3.2 matrix sentence 
This is based on the interpretation of the sentence. Thus, 1 polar_question and 2 wh_question categories include cases where it is a matrix polar/wh_question based on its sentence structure (e.g., A-not-A at the matrix level as in 觉不觉得 jue-bu-juede or 知不知道 zhi-bu-zhidao, or wh-words at the matrix level as in 怎么/为什么觉得/知道 zenme/weishenm juede/zhidao "how/why does [someone] think/know) and cases where the sentence has a declarative sentence structure ("superficially declarative") but contains a polar-construction/wh-word in the embedded clause that takes a matrix scope (which is sometimes referred to as "potentially interrogatives").

```{r matrix sentence type summary, message=FALSE}
matrix_sentence.data <- clean_data %>% 
  select(c(id, verb, sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3)) %>% 
  rowwise() %>% 
  mutate(
    n_filled = sum(c_across(sentence_type_e:sentence_type_j3) != "" & !is.na(c_across(sentence_type_e:sentence_type_j3))), # cases where only 1 coder considers it incomplete and the other considers it complete 
    agreement = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      if_else(n_filled == 2 && length(unique(vals)) == 1, "same", "diff")
      },
    agreed_sentence_type = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      vals[1]
    }
    ) %>%
  ungroup() %>% 
  filter(agreement == "same") %>% 
  select(!c(agreement, sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3, n_filled)) %>% 
  rename(matrix_sentence_type = agreed_sentence_type)

matrix_sentence_summary <- matrix_sentence.data %>% 
  group_by(verb, matrix_sentence_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

```{r matrix sentence type plot}
ggplot(matrix_sentence_summary, 
       aes(x=verb,
           y=percent,
           fill=matrix_sentence_type)) + 
  geom_col()+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_brewer(palette = "Set3",
                    name="Matrix sentence type",
                    guide=guide_legend(nrow = 2)) +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```

## 3.3 embedded sentence
"3 base position" includes cases where the a matrix sentence is superifically declarative but contains a polar-construction/wh-word in the embedded clause that takes a matrix scope (which is sometimes referred to as "potentially interrogatives").

```{r embedded sentence type summary, message=FALSE}
embedded_sentence.data <- clean_data %>% 
  select(c(id, verb, embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3)) %>% 
  mutate(
    across(embedded_sentence_type_e:embedded_sentence_type_j3, ~ if_else(.x == "", "NONE", .x))
  ) %>% 
  rowwise() %>% 
  mutate(
    agreement = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(length(unique(vals)) == 1, "same", "diff")
    },
    agreed_embedded_sentence_type = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(agreement == "same", vals[1], NA_character_)
    }
  ) %>%
  ungroup() %>% 
  filter(agreement == "same") %>% 
  select(!c(agreement, embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3)) %>% 
  rename(embedded_sentence_type = agreed_embedded_sentence_type)

embedded_sentence_summary <- embedded_sentence.data %>% 
  group_by(verb, embedded_sentence_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

```{r embedded sentence type plot}
ggplot(embedded_sentence_summary, 
       aes(x=verb,
           y=percent,
           fill=embedded_sentence_type)) + 
  geom_col()+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_brewer(palette = "Set3",
                    name="Embedded CP type",
                    guide=guide_legend(nrow = 2)) +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```

### 3.3.1 matrix and embedded sentence combined
Here we combined the annotations for matrix sentence type and embedded sentence type to analyze cases that are "potentially interrogative," i.e., the matrix clause type is in either 1 polar_question or 2 wh_question category AND the embedded clause is labeled as 3 base_position.

```{r matrix and embedded sentence type combined, message=FALSE}
matrix_embedded_sentence.data <- clean_data %>% 
  select(c(id, verb, 
           sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3,
           embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3)) %>% 
  mutate(
    across(embedded_sentence_type_e:embedded_sentence_type_j3, ~ if_else(.x == "", "NONE", .x))
  ) %>% 
  rowwise() %>% 
  mutate(
    matrix_n_filled = sum(c_across(sentence_type_e:sentence_type_j3) != "" & !is.na(c_across(sentence_type_e:sentence_type_j3))),
    matrix_agreement = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      if_else(matrix_n_filled == 2 && length(unique(vals)) == 1, "same", "diff")
    },
    embedded_agreement = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(length(unique(vals)) == 1, "same", "diff")
    },
    agreed_sentence_type = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      vals[1]
    },
    agreed_embedded_sentence_type = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(embedded_agreement == "same", vals[1], NA_character_)
    }
  ) %>%
  ungroup() %>% 
  filter(matrix_agreement == "same" & embedded_agreement == "same") %>% 
  select(!c(matrix_n_filled, matrix_agreement,sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3,
            embedded_agreement, embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3)) %>% 
  rename(matrix_sentence_type = agreed_sentence_type,
         embedded_sentence_type = agreed_embedded_sentence_type)
```

```{r matrix sentence type in combined summary, message=FALSE}
matrix_embedded_sentence_combined.data <- matrix_embedded_sentence.data %>%
  mutate(matrix_sentence_surface_type =
           case_when(matrix_sentence_type=="2 wh_question" & embedded_sentence_type=="3 base_position" ~ "0 declarative",
                     matrix_sentence_type=="1 polar_question" & embedded_sentence_type=="3 base_position" ~ "0 declarative",
                     TRUE ~ matrix_sentence_type),
         embedded_sentence_surface_type =
           case_when(matrix_sentence_type=="2 wh_question" & embedded_sentence_type=="3 base_position" ~ "3 base_position: wh",
                     matrix_sentence_type=="1 polar_question" & embedded_sentence_type=="3 base_position" ~ "3 base_position: polar",
                     TRUE ~ embedded_sentence_type)
           )

matrix_combined_summary <- matrix_embedded_sentence_combined.data %>% 
  group_by(verb, matrix_sentence_surface_type,matrix_sentence_type) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

The greenish/teal area includes cases where the sentence that has a declarative sentence stucture (i.e., "superficially declarative"). Within the greenish area, the dotted part represents the sentence with a wh-word in the embedded clause that takes a matrix scope, such that the superficially declarative sentence is interpreted as a wh question. Likewise, the striped part in the greenish area represents sentences with a polar construction (e.g., A-not-A or *haishi*) in the embedded clause that takes a matrix scope, such that the superficially declarative sentence is interpreted as a polar question.


```{r matrix sentence type plot in combined summary}
ggplot(matrix_combined_summary, 
       aes(x=verb,
           y=percent,
           fill=matrix_sentence_surface_type,
           pattern=matrix_sentence_type)) + 
  geom_col_pattern(
    color = "black",
    linewidth = 0.2,
    pattern_fill = "lightgrey",
    pattern_density = 0.1,
    pattern_size = 0.2,
    pattern_spacing = 0.03,
    pattern_alpha = 0.4
  )+
  scale_pattern_manual(
    values = c(
      "0 declarative" = "none",
      "1 polar_question" = "stripe",
      "2 wh_question" = "circle",
      "4 tag_question" = "none"),
    name="Matrix sentence type",
    guide=guide_legend(nrow=2)
    )+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_brewer(palette = "Set3",
                    name="Matrix sentence type",
                    guide=guide_legend(nrow=2)) +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```
```{r embedded sentence type in combined summary, message=FALSE}
embedded_combined_summary <- matrix_embedded_sentence_combined.data %>% 
  group_by(verb, embedded_sentence_surface_type) %>%
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

```{r embedded sentence type plot in combined summary}
ggplot(embedded_combined_summary %>% 
         mutate(embedded_sentence_surface_type = fct_relevel(
           embedded_sentence_surface_type,
           "0 declarative",
           "1 polar_question",
           "3 base_position: polar",
           "2 wh_question",
           "3 base_position: wh",
           "NONE"
         )), 
       aes(x=verb,
           y=percent,
           fill=embedded_sentence_surface_type,
           pattern=embedded_sentence_surface_type)) + 
  geom_col_pattern(
    color = "black",
    linewidth = 0.2,
    pattern_fill = "lightgrey",
    pattern_density = 0.1,
    pattern_size = 0.2,
    pattern_spacing = 0.03,
    pattern_alpha = 0.4
  )+
  scale_pattern_manual(
    values = c(
      "0 declarative" = "none",
      "1 polar_question" = "none",
      "2 wh_question" = "none",
      "3 base_position: polar" = "circle",
      "3 base_position: wh" = "circle",
      "NONE" = "none"),
    name="Embedded CP type",
    guide=guide_legend(nrow=2)
  )+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_manual(
    values = c(
      "0 declarative" = "#8DD3C7",  # 1
      "1 polar_question" = "#FFFFB3",  # 2
      "3 base_position: polar" = "#FFFFB3",
      "2 wh_question" = "#BEBADA", # 3
      "3 base_position: wh" = "#BEBADA",
      "NONE" = "#80B1D3" # 5
    ),
    name = "Embedded CP type",
    guide=guide_legend(nrow=2)
  )+
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```

## 3.4 matrix negation
```{r matrix negation summary, message=FALSE}
matrix_negation.data <- clean_data %>% 
  select(c(id, verb, matrix_negation_e, matrix_negation_j1, matrix_negation_g)) %>% 
  rowwise() %>% 
  mutate(
    agreement = {
      vals <- c_across(matrix_negation_e:matrix_negation_g)
      vals <- vals[!is.na(vals)]
      if_else(length(unique(vals)) == 1, "same", "diff")
    },
    agreed_negation_type = {
      vals <- c_across(matrix_negation_e:matrix_negation_g)
      vals <- vals[!is.na(vals)]
      if_else(agreement == "same", vals[1], NA_character_)
    }
  ) %>%
  ungroup() %>% 
  filter(agreement == "same") %>% 
  select(!c(agreement, matrix_negation_e, matrix_negation_j1, matrix_negation_g)) %>% 
  rename(negation_type = agreed_negation_type)

matrix_negation_summary <- matrix_negation.data %>% 
  mutate(negation = if_else(negation_type == "0 no negation", negation_type, "1 negation")) %>% 
  group_by(verb, negation) %>% 
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()
```

```{r negaton plot}
ggplot(matrix_negation_summary, 
       aes(x=verb,
           y=percent,
           fill=negation)) + 
  geom_col()+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_brewer(palette = "Set3",
                    name="Matrix negation type") +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")

```
## 3.5 speech acts
We are using syntactic structures as proxies for different speech acts. Only data that annotators agreed on all categories (i.e., matrix subject type, matrix and embedded sentence types, and negation) are included.
```{r speech acts data}
subject_sentence.data <- clean_data %>% 
  select(c(id, verb,
           subject_j1, subject_j2,
           sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3,
           embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3,
           matrix_negation_e, matrix_negation_j1, matrix_negation_g)) %>% 
  mutate(subject_agreement = if_else(subject_j1 == subject_j2, "same", "diff"),
    across(embedded_sentence_type_e:embedded_sentence_type_j3, ~ if_else(.x == "", "NONE", .x))
  ) %>% 
  filter(subject_agreement=="same") %>% 
  select(!c(subject_agreement, subject_j1)) %>%
  rowwise() %>% 
  mutate(
    matrix_n_filled = sum(c_across(sentence_type_e:sentence_type_j3) != "" & !is.na(c_across(sentence_type_e:sentence_type_j3))),
    matrix_agreement = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      if_else(matrix_n_filled == 2 && length(unique(vals)) == 1, "same", "diff")
    },
    embedded_agreement = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(length(unique(vals)) == 1, "same", "diff")
    },
    negation_agreement = {
      vals <- c_across(matrix_negation_e:matrix_negation_g)
      vals <- vals[!is.na(vals)]
      if_else(length(unique(vals)) == 1, "same", "diff")
    },
    agreed_sentence_type = {
      vals <- c_across(sentence_type_e:sentence_type_j3)
      vals <- vals[vals != "" & !is.na(vals)]
      vals[1]
    },
    agreed_embedded_sentence_type = {
      vals <- c_across(embedded_sentence_type_e:embedded_sentence_type_j3)
      vals <- vals[!is.na(vals)]
      if_else(embedded_agreement == "same", vals[1], NA_character_)
    },
    agreed_negation_type = {
      vals <- c_across(matrix_negation_e:matrix_negation_g)
      vals <- vals[!is.na(vals)]
      if_else(negation_agreement == "same", vals[1], NA_character_)
    }
  ) %>%
  ungroup() %>% 
  filter(matrix_agreement == "same" & embedded_agreement == "same" & negation_agreement == "same") %>% 
  select(!c(matrix_n_filled, matrix_agreement,sentence_type_e, sentence_type_g, sentence_type_c, sentence_type_j3,
            embedded_agreement, embedded_sentence_type_e, embedded_sentence_type_g, embedded_sentence_type_c, embedded_sentence_type_j3,
            matrix_negation_e, matrix_negation_j1, matrix_negation_g,
            negation_agreement)) %>% 
  rename(subject_type = subject_j2,
         matrix_sentence_type = agreed_sentence_type,
         embedded_sentence_type = agreed_embedded_sentence_type,
         negation_type = agreed_negation_type)
```

Here are the syntactic combinations that we are considering:

- **(I) don't know/think**: first person or dropped subject, matrix declarative, no embedded, negation
  - e.g., 我不知道 I don't know. 不知道 Don't know
  
- **I know/think P**: first person subject, matrix declarative, embedded declarative, no negation
  - e.g., 我觉得该睡觉了 I think it is time for bed. 我知道它是一只猫 I know it is a cat.
  
- **You (don't) know/think P**: second person subject, matrix declarative, embedded declarative, no/yes negation
  - e.g., 你觉得该睡觉了 You think it is time for bed. 你知道它是一只猫 You know it is a cat.
  
- **do you know/think WH**: second person subject, matrix polar, embedded wh, no negation
  - e.g., 你知道他是谁吗？Do you know who he is? (structurally: You know he is who?)
  
- **WH do you know/think**: base : second person subject, matrix wh, embedded base position, no negation
  - e.g., 你觉得他是谁？Who do you think he is? (structurally: You think he is who?)
  
- **do you know/think P**: second person subject, matrix polar, embedded declarative, no negation
  - e.g., 你觉得该睡觉了吗？Do you think it is time for bed? 你知道它是一只猫吗？Do you know it is a cat? 
  
- **do you know/think P**: base: second person subject, matrix polar, embedded base position, no negation
  - e.g., 你觉得是不是该睡觉了？Do you think it is time for bed or not? (structurally: You think it is or it is not time for bed)
  
- **(you) know/think**: second person or dropped subject, matrix polar, no embedded, no negation
  - e.g., 你觉得吗？Do you think? 知不知道？Do (you) know?
  
- **(I) know/think**: first person or dropped subject, matrix declarative, no embedded, no negation
  - e.g., 我知道 I know. 知道 Know
  
- **tag question**: matrix tag question
  - e.g., 今天下雨了，知道吗？It rained today, you know? 你知道今天下雨了，对吧？You knew it rained today, right?
  
- **third person**: matrix subject is a third person pronoun or an NP (regardless of sentence type and negation)
  - e.g., 他觉得今天下雨了 He thought it rained today 他不知道今天下雨了吗？ Did he not know it rained today?

```{r speech act summary, message=FALSE}
speech_acts.data <- subject_sentence.data %>% 
  mutate(negation = if_else(negation_type == "0 no negation", "0 no negation", "1 yes negation")) %>% 
  mutate(speech_act = case_when(
    subject_type %in% c("1 first_person", "0 dropped") & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "NONE" & negation == "1 yes negation" ~ "(i) don't know/think",
    subject_type %in% c("1 first_person") & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "0 declarative" & negation == "0 no negation" ~ "i think/know P",
    subject_type == "2 second_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "0 declarative" ~ "you (don't) think/know P",
    subject_type == "2 second_person" & matrix_sentence_type == "1 polar_question" & embedded_sentence_type == "2 wh_question" & negation == "0 no negation" ~ "do you know WH",
    subject_type == "2 second_person" & matrix_sentence_type == "2 wh_question" & embedded_sentence_type == "3 base_position" & negation == "0 no negation" ~ "WH do you think/know: base",
    subject_type == "2 second_person" & matrix_sentence_type == "1 polar_question" & embedded_sentence_type == "0 declarative" & negation == "0 no negation" ~ "do you know/think P",
    subject_type == "2 second_person" & matrix_sentence_type == "1 polar_question" & embedded_sentence_type == "3 base_position" & negation == "0 no negation" ~ "do you think/know P: base",
    subject_type %in% c("0 dropped", "2 second_person") & matrix_sentence_type == "1 polar_question" & embedded_sentence_type == "NONE" & negation == "0 no negation" ~ "(you) know/think?",
    subject_type %in% c("1 first_person", "0 dropped") & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "NONE" & negation == "0 no negation" ~ "(i) know/think",
    matrix_sentence_type == "4 tag_question" ~ "tag question",
    subject_type == "3 third/NP" ~ "third person",
    TRUE ~ "others"))

speech_acts_summary <- speech_acts.data %>% 
  group_by(verb, speech_act) %>%
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup() %>% 
  mutate(speech_act=fct_relevel(speech_act,
                                "do you know WH",
                                "WH do you think/know: base",
                                "do you think/know P: base",
                                "do you know/think P",
                                "i think/know P",
                                "you (don't) think/know P",
                                "(i) don't know/think",
                                "(i) know/think",
                                "(you) know/think?",
                                "tag question",
                                "third person",
                                "others"))
```

```{r speech act plot}
ggplot(speech_acts_summary, 
       aes(x=verb,
           y=percent,
           fill=speech_act,
           pattern=speech_act)) + 
  geom_col_pattern(
    color = "black",
    linewidth = 0.2,
    pattern_fill = "lightgrey",
    pattern_density = 0.1,
    pattern_size = 0.2,
    pattern_spacing = 0.03,
    pattern_alpha = 0.6
  )+
  scale_pattern_manual(
    values = c(
      "do you know WH" = "none",
      "WH do you think/know: base" = "circle",
      "do you know/think P" = "none",
      "do you think/know P: base" = "stripe",
      "you (don't) think/know P" = "none",
      "i think/know P" = "none",
      "(i) don't know/think" = "none",
      "(i) know/think" = "none",
      "(you) know/think?" = "none",
      "tag question" = "none",
      "third person" = "none",
      "others" = "none"),
    name="Speech acts type",
    guide=guide_legend(nrow=4)
  )+
  geom_text(aes(label=paste0(count, " (", round(percent, 1), "%)")),
            position = position_stack(vjust = 0.5),
            color = "black",
            size = 3)+
  theme_bw()+
  scale_fill_manual(
    values = c(
      "do you know WH" = "#CC79A7", # purple
      "WH do you think/know: base" = "#CC79A7",
      "do you know/think P" = "#56B4E9",  # blue # light blue #6E8894
      "do you think/know P: base" = "#56B4E9",
      "i think/know P" = "#E69F00", # orange
      "you (don't) think/know P"= "#E2C391",
      "(i) don't know/think" = "#009E73", # green
      "(i) know/think" = "#B4CEB3",
      "(you) know/think?" = "#A4DDF8", # light yellow #F6E27F
      "tag question" = "#D55E00",
      "third person" = "#F0E442", # yellow
      "others" = "grey80"
    ),
    name="Speech acts type",
    guide=guide_legend(nrow=4)
  ) +
  theme(legend.position = "top",
        axis.title.x = element_text(size = 14),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(size = 12),
        strip.text.x = element_text(size=12),
        legend.text = element_text(size=10),
        legend.title = element_text(size=12),
        axis.title.y = element_text(size = 14))+
  labs(x="Verb",
       y="Percentage")
```

```{r other speech acts, message=FALSE}
other_speech_acts.data <- speech_acts.data %>% 
  filter(speech_act=="others") %>% 
  mutate(speech_act = case_when(
    subject_type == "2 second_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type %in% c("1 polar_qustion", "2 wh_question") & negation == "1 yes negation" ~ "you don't think/know P/WH",
    subject_type == "2 second_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type %in% c("1 polar_qustion", "2 wh_question") & negation == "0 no negation" ~ "you think/know P/WH",
    subject_type == "2 second_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "NONE" & negation == "1 yes negation" ~ "you don't think/know",
    subject_type == "2 second_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "2 wh_question" & negation == "0 no negation" ~ "you think/know WH",
    subject_type == "0 dropped" & matrix_sentence_type == "0 declarative" & embedded_sentence_type == "2 wh_question" ~ "(don't) think/know WH",
    subject_type == "1 first_person" & matrix_sentence_type == "0 declarative" & embedded_sentence_type %in% c("1 polar_question","2 wh_question") ~ "i (don't) think/know (P/WH)",
    subject_type == "3 third/NP" & matrix_sentence_type == "0 declarative" & embedded_sentence_type %in% c("1 polar_qustion", "2 wh_question", "NONE") ~ "she/he/NP (don't) think/know (P/WH)",
    subject_type == "3 third/NP" & matrix_sentence_type %in% c("1 polar_qustion", "2 wh_question") & embedded_sentence_type %in% c("0 declarative", "NONE") ~ "does/WH she/NP (don't) think/know (P)",
    TRUE ~ "others"))

other_speech_acts_summary <- other_speech_acts.data %>% 
  group_by(verb, speech_act) %>%
  summarize(count = n()) %>% 
  ungroup() %>% 
  group_by(verb) %>% 
  mutate(percent = count/sum(count) * 100) %>% 
  ungroup()

other_speech_acts_summary
```

